Weighted Visibly Pushdown Automata and Automated Music Transcription

### 2021-05-01 main.tex
1-best algo

%% theorem envs
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

@@

@@

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\W}{\mathbb{W}}

@@

\def\SWT{\textsf{SWT}\xspace}
\def\SWA{\textsf{SWA}\xspace}
\def\SWVPA{\textsf{SWVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}}

@@

@@

\noindent
A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
%
Following the terminology of~\cite{Mohri02semiring},
when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
the semiring $\Semiring$ is is called \emph{bounded}.
Note that every bounded semiring is idempotent:
by boundedness, 
$\one \oplus \one = \one$, and idempotency follows by multiplying
both sides by $x$ and distributing.

@@

Every idempotent semiring~$\Semiring$ induces 
a partial ordering~$\leq_\Semiring$ 
called the \emph{natural ordering} of~$\Semiring$
and defined by: 
%implicitly defined by the semiring $\Semiring$ 
for all $x$ and $y$,
$x \leq_\Semiring y \;\mbox{iff}\; x \oplus y = x$.
This ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
The above definition follows \cite{Mohri02semiring}, 
and coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
%
It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\Semiring$.
An idempotent semiring $\Semiring$~is called \emph{total} if
it~$\leq_\Semiring$ is total
\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.

@@

@@

\noindent 
A \emph{label theory} is made of 4 recursively enumerable sets:
%$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
$\Phi_\Sigma$ and $\Phi_\Delta$, 
containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
Moreover, we assume that these sets are closed under the operators 
$\oplus$ and $\otimes$ of~$\Semiring$.
More precisely, for all $\phi, \phi' \in \Phi_\Sigma$ all 
$\psi, \psi' \in \Phi_\Delta$, 
and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
\begin{description}
\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
\end{description}
The same also holds for the binary sum operator $\oplus$.

\noindent
\marginpar{maybe not necessary}
Finally, it is assumed that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
is a subset of $\Phi_\epsilon$.
and all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
belong resp. to $\Phi_\Sigma$ and $\Phi_\Delta$.

%(SWT)
\begin{definition}
A \emph{symbolic-weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
and the semiring $\Semiring$ is a tuple
$T = \< Q, \init, \wei, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\wei$ is a transition %weight 
function from $Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
\end{definition}
%
We extend the above transition function into a function from
$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
into $\Semiring$, also called $\wei$ for simplicity, such that 
for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
and with 
$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
\[
\begin{array}{rcl}
\wei(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
\wei(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
\end{array}      
\]
%More precisely, $Q \times Q$, 
%resp. $Q \times \Sigma \times Q$,
%$Q \times \Delta \times Q$,
%$Q \times \Sigma \times \Delta \times Q$,
%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.

@@

The symbolic-weighted transducer $T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into the weights of~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, s, t, q') & = & & \wei(q, \epsilon, \epsilon, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                      \scriptstyle q'' \in Q\\[-2pt]
                                      \scriptstyle s = au, a \in \Sigma
                                      \end{array}} &
     \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
      \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                         \scriptstyle q'' \in Q\\[-2pt]
                                         \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                         \scriptstyle t = bv, b \in \Delta
                                         \end{array}} &
      \wei(q, a, b, q'') \otimes \weight_\A(q'', u, v, q')\\ 
\end{array}
\]
Recall that by convention, an empty sum with $\oplus$ is $\zero$. 
%
\noindent
The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
is then defined as follows: 
\[
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
\]

A \emph{symbolic weighted automata} (\SWA) $A = \< Q, \init, \weight, \final >$
over $\Sigma$ and $\Semiring$ 
is defined in a similar way by simply omitting the output symbols,
\ie $\wei$ is a function of~$Q \times Q$ into %$\Phi$.
$\< \Phi_0, \Phi_\Sigma >$, 
or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.

\begin{proposition}
Given a \SWT $T$ 
over $\Sigma$, $\Delta$ and $\Semiring$, 
and a word $s \in \Sigma^*$, 
one can construct a \SWA 
$A_{s, T}$ such that for all $t \in \Delta^*$, 
$A_{s, T}(t) = T(s, t)$.
\end{proposition}
The construction time and size of $A_{s, T}$ are $O(| s | . \| T \|)$.

@@

Let $\Sigma$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into :
\begin{itemize}
\item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
\item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
\item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
\end{itemize}
Let $\Semiring$ be a commutative semiring and  
let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$.
%
Moreover, we extend this theory with a set $\Phii$ 
of unary functions in $\Sigmai \to \Semiring$,
closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
over  $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$ and $\Semiring$ is a tuple
$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
where $Q$ is a finite set of states, 
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and 
$\weii : Q \times Q \to \Phii$,  
$\weic : Q \times Q \times P \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weie : Q \times Q \to \Phir$,  
are transition functions.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Sigmai$, 
$\call{c} \in \Sigmac$, 
$\return{r} \in \Sigmar$, 
overloading the names for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Sigmai \times Q \to \Semiring & 
\weii(q, a, q') = \phi_\mathsf{i}(a) & 
\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
%
\weic: Q \times \Sigmac \times Q \times P \to \Semiring & 
\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
%
\weir: Q \times \Sigmac \times P \times \Sigmar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
%
\weie: Q \times \Sigmar \times Q \to \Semiring & 
\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
\end{array}      
\]

\noindent
The intuition is the following for the above transitions.
\begin{description}
\item $\weii$ : read the input internal symbol $a$, change state to $q'$.
\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
\item $\weir$ : when the stack is not empty, 
      read and pop from stack a pair made of $\call{c}$ and $p$, 
      read the input return symbol $\return{r}$, change state to $q'$.
      In this case, the weight function $\phi_\mathsf{r}$ 
      computes the value of matching between the call and return symbols.
      It might be $\zero$ to express that the symbols do not match.
\item $\weie$ : when the stack is empty, 
      read the input symbol $\call{r}$, change state to $q'$.
\end{description}

We give now a formal definition of these transitions of the automaton $A$
in term of a weight value
computed by an intermediate function $\weight_A$.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Sigmac \times P$.
Therefore, $\weight_A$ is a function from 
$Q \times \Theta^* \times \Sigma^* \times Q \times \Theta^*$ into $\Semiring$.
\[
\begin{array}{rcl}
\weight_A\bigl(\config{q}{\theta}, a u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\theta}, {\call{c}} u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, q'', p\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}} u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
\end{array}
\]
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
denotes a stack with the pair made of ${\call{c}}$ and $p$ on its top and $\theta$
as the rest of stack.

@@

\subsection{1-best Computation}

Regarding the infinite sum operator, it must hold that
$\bigoplus_{x \in \Phi_\Sigma} \phi(x)$, 
$\bigoplus_{y \in \Phi_\Delta} \psi(y)$, and

>>>>

@@

%% theorem envs
\usepackage{theorem}
\newtheorem{theorem}{Theorem} %[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
\theorembodyfont{\slshape}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

%% algorithms
%\usepackage{algorithm} 
%\usepackage{program}

@@

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\W}{\mathbb{W}}

@@

\def\SWT{\textsf{SWT}\xspace}
\def\SWA{\textsf{SWA}\xspace}
\def\SWVPA{\textsf{SWVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}

@@

weighted parsing

principle of quantitative parsing / parsing of words over infinite alphabet:

@@

@@

%\noindent
A semiring $\Semiring$ is \emph{commutative} if $\otimes$ is commutative.
It is \emph{idempotent} if for each $x \in \dom(\Semiring)$, $x \oplus x = x$.
%
%Following the terminology of~\cite{Mohri02semiring},
%when $\forall x \in \dom(\Semiring), \one \oplus x = \one$,
%the semiring $\Semiring$ is is called \emph{bounded}.
%Note that every bounded semiring is idempotent:
%by boundedness, 
%$\one \oplus \one = \one$, and idempotency follows by multiplying
%both sides by $x$ and distributing. 
%
Every idempotent semiring~$\Semiring$ induces 
a partial ordering~$\leq_\oplus$ 
called the \emph{natural ordering} of~$\Semiring$
and defined by: 
%implicitly defined by the semiring $\Semiring$ 
for all $x$ and $y$,
$x \leq_\oplus y \;\mbox{iff}\; x \oplus y = x$.
This ordering is sometimes defined in the opposite direction~\cite{DrosteKuich09semirings};
The above definition follows \cite{Mohri02semiring}, 
and coincides than the usual ordering on the Tropical semiring (\emph{min-plus}).
%
It holds that $\Semiring$ is {monotonic} \wrt~$\leq_\oplus$.
An idempotent semiring $\Semiring$~is called \emph{total} if
it~$\leq_\oplus$ is total
\ie when for all $x, y \in \Semiring$, either $x \oplus y = x$ or $x \oplus y = y$.

@@

\begin{example}
semirings      
\end{example}

@@

\noindent 
A \emph{label theory} over $\Sigma$ and $\Delta$
is made of 4 recursively enumerable sets:
%$\Phi_0$, \Phi_\Sigma \uplus \Phi_\Delta \uplus \Phi_{\Sigma, \Delta}$ 
$\Phi_\epsilon \subseteq \Semiring$, % containing constant functions valued in $\Semiring$, 
$\Phi_\Sigma$ and $\Phi_\Delta$, 
containing unary functions in $\Sigma \to \Semiring$, resp. $\Delta \to \Semiring$, 
and $\Phi_{\Sigma, \Delta}$  containing binary functions in $\Sigma \times \Delta \to \Semiring$.
Moreover, we assume that these sets are closed under the operators 
$\oplus$ and $\otimes$ of~$\Semiring$.
More precisely, for all $\phi, \phi' \in \Phi_\Sigma$ all 
$\psi, \psi' \in \Phi_\Delta$, 
and $\eta, \eta' \in \Phi_{\Sigma, \Delta}$, the function
\begin{description}
\item $\phi \otimes \phi' : x \mapsto \phi(x) \otimes \phi'(x)$ belongs to $\Phi_\Sigma$, 
\item $\psi \otimes \psi' : y \mapsto \psi(y) \otimes \psi'(y)$ belongs to $\Phi_\Delta$,
\item $\phi \otimes \eta : x, y \mapsto \phi(x) \otimes \eta(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$,
\item $\eta \otimes \psi : x, y \mapsto \eta(x, y) \otimes \psi(y)$ belongs to $\Phi_{\Sigma, \Delta}$,
\item $\eta \otimes \eta' : x, y \mapsto \eta(x, y) \otimes \eta'(x, y)$ belongs to $\Phi_{\Sigma, \Delta}$.
\end{description}
The same also holds for the binary sum operator $\oplus$.

\noindent
\marginpar{necessary?}
Finally, it is assumed that the codomain of every function of $\Phi_\Sigma$ and $\Phi_\Delta$
is a subset of $\Phi_\epsilon$.
and all partial applications of functions $\Phi_{\Sigma, \Delta}$, 
resp.  $f_a: y \mapsto f(a, y)$ for $a \in \Sigma$ and $y \in \Delta$
and  $f_b: x \mapsto f(x, b)$ for $b \in \Delta$ and $x \in \Sigma$, 
belong resp. to $\Phi_\Sigma$ and $\Phi_\Delta$.

\subsection{Definitions}
%(SWT)
\begin{definition}
A \emph{symbolic-weighted transducer} $T$ over the input and output alphabet $\Sigma$ and $\Delta$ 
and the semiring $\Semiring$ is a tuple
$T = \< Q, \init, \wei, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\wei$ is a transition %weight 
function from $Q \times Q$ into %$\Phi$.
$\< \Phi_\epsilon, \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$.
\end{definition}
%
We extend the above transition function into a function from
$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q$
into $\Semiring$, also called $\wei$ for simplicity, such that 
for all $q, q' \in Q$, $a \in \Sigma$, $b \in \Delta$, 
and with 
$\< \phi_\epsilon, \phi_\Sigma, \phi_\Delta, \phi_{\Sigma, \Delta}> = \wei(q, q')$, 
\[
\begin{array}{rcl}
\wei(q, \epsilon, \epsilon, q') & = & \phi_\epsilon\\
\wei(q, a, \epsilon, q') & = & \phi_\Sigma(a)\\
\wei(q, \epsilon, b, q') & = & \phi_\Delta(b)\\
\wei(q, a, b, q') & = & \phi_{\Sigma, \Delta}(a, b)
\end{array}      
\]
%More precisely, $Q \times Q$, 
%resp. $Q \times \Sigma \times Q$,
%$Q \times \Delta \times Q$,
%$Q \times \Sigma \times \Delta \times Q$,
%into resp. $\Phi_0$, $\Phi_\Sigma$, $\Phi_\Delta$ $\Phi_{\Sigma, \Delta}$.

@@

The symbolic-weighted transducer $T$ defines a mapping 
from the pairs of strings of $\Sigma^* \times \Delta^*$ 
into the weights of~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
\[
\begin{array}{rccl}
\weight_T(q, s, t, q') & = & & \wei(q, \epsilon, \epsilon, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                      \scriptstyle q'' \in Q\\[-2pt]
                                      \scriptstyle s = au, a \in \Sigma
                                      \end{array}} &
     \wei(q, a, \epsilon, q'') \otimes \weight_T(q'', u, t, q')\\
 & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                    \scriptstyle q'' \in Q\\[-2pt]
                                    \scriptstyle t = bv, b \in \Delta
                                    \end{array}} &
      \wei(q, \epsilon, b, q'') \otimes \weight_T(q'', s, v, q')\\ 
      & & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                         \scriptstyle q'' \in Q\\[-2pt]
                                         \scriptstyle s = au, a \in \Sigma\\[-2pt]
                                         \scriptstyle t = bv, b \in \Delta
                                         \end{array}} &
      \wei(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\\ 
\end{array}
\]
Recall that by convention, an empty sum with $\oplus$ is $\zero$. 
%
\noindent
The weight associated by $T$ to  $\< s, t> \in \Sigma^* \times \Delta^*$
is then defined as follows: 
\[
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q').
\]

\begin{example}
comparison of two sequences of timestamped events      
\end{example}

\noindent
A \emph{symbolic weighted automata} (\SWA) $A = \< Q, \init, \weight, \final >$
over $\Sigma$ and $\Semiring$ 
is defined in a similar way by simply omitting the output symbols,
\ie $\wei$ is a function from~$Q \times Q$ into %$\Phi$.
$\< \Phi_\epsilon, \Phi_\Sigma >$, 
or equivalently from~$Q \times (\Sigma \cup \{ \epsilon \}) \times Q$ into~$\Semiring$.

\subsection{Properties}

\begin{proposition}
Given a \SWT $T$ 
over $\Sigma$, $\Delta$ and $\Semiring$, 
and a word $s \in \Sigma^*$, 
one can construct a \SWA 
$A_{s, T}$ such that for all $t \in \Delta^*$, 
$A_{s, T}(t) = T(s, t)$.
\end{proposition}
The construction time and size for $A_{s, T}$ are $O(| s | . \| T \|)$.

\begin{proposition}
      ...removal of $\epsilon$ transitions for \SWA...
      \end{proposition}

@@

\subsection{Definition}
Let $\Sigma$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three subsets 
$\Sigmai \uplus \Sigmac \uplus \Sigmac$ 
respectively called of \emph{internal}, \emph{call} and \emph{return}
symbols.
% \begin{itemize}
% \item a set $\Sigmai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Sigmac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Sigmar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let $\Semiring$ be a commutative semiring and  
let  $(\Phi_\epsilon, \Phic, \Phir, \Phicr)$ be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phic$, $\Phir$ and~$\Phicr$ stand respectively 
for~$\Phi_\Sigmac$, $\Phi_\Sigmar$ and~$\Phi_{\Sigmac, \Sigmar}$.
%
Moreover, we extend this theory with a set $\Phii$ 
of unary functions in $\Sigmai \to \Semiring$,
closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)~$A$ 
over  $\Sigma = \Sigmai \uplus \Sigmac \uplus \Sigmar$ and $\Semiring$ is a tuple
$T = \< Q, P, \init, \weii, \weic, \weir, \weie, \final >$,
where $Q$ is a finite set of states, 
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$, 
respectively $\mathsf{out} : Q \to \Semiring,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and 
$\weii : Q \times Q \to \Phii$,  
$\weic : Q \times Q \times P \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weie : Q \times Q \to \Phir$,  
are transition functions.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Sigmai$, 
$\call{c} \in \Sigmac$, 
$\return{r} \in \Sigmar$, 
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Sigmai \times Q \to \Semiring & 
\weii(q, a, q') = \phi_\mathsf{i}(a) & 
\mathrm{where~} \phi_\mathsf{i} = \weii(q, q'),\\
%
\weic: Q \times \Sigmac \times Q \times P \to \Semiring & 
\weic(q, \call{c}, q', p) = \phi_\mathsf{c}(\call{c}) & 
\mathrm{where~} \phi_\mathsf{c} = \weic(q, q', p),\\
%
\weir: Q \times \Sigmac \times P \times \Sigmar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \phi_\mathsf{r}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \phi_\mathsf{r} = \weir(q, p, q'),\\
%
\weie: Q \times \Sigmar \times Q \to \Semiring & 
\weie(q, {\return{r}}, q') = \phi_\mathsf{e}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{e} = \weie(q, q').\\
\end{array}      
\]

\noindent
The intuition is the following for the above transitions.
\begin{description}
\item $\weii$ : read the input internal symbol $a$, change state to $q'$.
\item $\weic$ : read the input call symbol $\call{c}$, push it to the stack along with $p$, change state to $q'$.
\item $\weir$ : when the stack is not empty, 
      read and pop from stack a pair made of $\call{c}$ and $p$, 
      read the input return symbol $\return{r}$, change state to $q'$.
      In this case, the weight function $\phi_\mathsf{r}$ 
      computes a value of matching between the call and return symbols.
      This value might be $\zero$ in order to express that the symbols do not match.
\item $\weie$ : when the stack is empty, 
      read the input symbol $\call{r}$, change state to $q'$.
\end{description}

We give now a formal definition of these transitions of the automaton $A$
in term of a weight value
computed by an intermediate function $\weight_A$.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\theta \in \Theta^*$, where $\Theta = \Sigmac \times P$.
Therefore, $\weight_A$ is a function from 
$Q \times \Theta^* \times \Sigma^* \times Q \times \Theta^*$ into $\Semiring$.
\[
\begin{array}{rcl}
\weight_A\bigl(\config{q}{\theta}, a\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\theta}, {\call{c}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, q'', p\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p\cdot \theta}, u, \config{q'}{\theta'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \theta}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\theta}, u, \config{q'}{\theta'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\theta'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weie(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\theta'}\bigr)\\
\end{array}
\]
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \theta$ 
denotes a stack with the pair made of ${\call{c}}$ and $p$ on its top and $\theta$
as the rest of stack.

@@

\begin{example}
structured words...
intro language of music notation ?
\end{example}
